\documentclass{article}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumerate}

\newtheorem*{thm}{Theorem}
\newtheorem*{cly}{Corollary}
\newtheorem*{prf}{Proof}
\newtheorem*{defi}{Definition}

\newcommand{\floor}[1]{\lfloor {#1} \rfloor}
\newcommand{\abs}[1]{\left| {#1} \right|}

\newcommand{\diag}{\mathop{\mathrm{diag}}}
\newcommand{\trace}{\mathop{\mathrm{trace}}}


% \title{}
% \author{Lgarithm}
\begin{document}
% \maketitle
% \section{}
% \subsection{}

Recall the union, contatenation, of two regular language is regular.
And the closure and complement of a regular language is regular.
so by De morgan's law, the intersection of two regular language is regular.

Let $R$ be an infinte regular language, if $R = G^\ast$ for some finite $G$, 
we say $R$ is finite generated, by $G$.

The set of all regular expressions, $RegExp$, can be viewed as a formal 
language over $\Sigma \cup S$, where $S$ is a finite set of 
symbols which may occur in a regular expression.

The definition of regular expression is recursive: 
\begin{enumerate}[i).]
\item $\Sigma \cup \{\epsilon, \emptyset\} \subseteq RegExp$;
\item if $r, g \in RegExp$, so is $(r \circ g)$, $(r \cup g)$, $(r^\ast)$;
\end{enumerate}

% From the definition we can see $RegExp$ is a context free language itself.

For a subset $G$ of $RegExp$, there is a subset $R$ of $\Sigma^\ast$, 
whose element matches every regular expression in $G$. 
And of course we know $R$ must be a regular language.

And for a subset $L$ of $\Sigma^\ast$, $L$ may not be regular, 
there is a subset $H$ of $RegExp$,
which contains all regular expressions that matches every words in $L$.

Problem: How to calculate the intersection of two regular language?The languages are represented by regular expression.
How to determine if the intersection is empty?

For two alphabet $\Sigma_1, \Sigma_2$ and a language $L$ over $\Sigma_1$, 
a semantic of $L$ is a map $\Sigma_1 \to \Sigma_2$ restrict to $L$.Two words of $L$ has the same semantic means their images are equal in $\Sigma_2$.

\end{document}
